---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(QPress)
library(tidyverse)
library(patchwork)
theme_set(theme_classic())
source('models.R')
source('helpers.R')
```

# How to build a model 
### Mangrove response to climate change and other anthropogenic threats

Taking a 'one model to rule them all approach' vs. multiple, individual models for different scenarios. We agree there should be only one model structure. If we think the model structure isn't right though, we can test and compare other variations.

Aim is to build one model that performs as expected under different scenarios.

Then we can map likelihood of increase or decrease in mangrove extent, given different drivers

Steps are to:

1) Build and visualise the conceptual model
2) Perturb drivers and simulate to examine expected responses of mangroves under different scenarios
3) If responses are not expected, re-work the conceptual model and do step 2
4) Map likelihood of mangroves increasing or decreasing given different geomorphologies and drivers
5) Validate with GMW data (if possible)

### 1. Build and visualise conceptual model

Variables are in white and drivers are in grey. Red arrows indicate negative effect, black arrows indicate positive effect.

```{r dpi=300, fig.width=7, fig.height=7}
plotmodel <- function(model, name){
w <- adjacency.matrix(model, labels=TRUE)
gdat <- graph_from_adjacency_matrix(t(w),
                                    mode = "directed",
                                    weighted = "b")
E(gdat)$color <- ifelse(E(gdat)$b <0, "red", "black")
#V(gdat)$color <- ifelse(names(V(gdat)) %in% drivers, "grey", "white")
plot(gdat, 
     vertex.size = 10,
     vertex.label.cex = 0.5,
     vertex.label.color = "black",
     edge.arrow.size = 0.3,
     #edge.label = letters[1:length(E(gdat))],
     edge.label.cex = 0.6,
     edge.label.color = "black",
     edge.label.pos = 1,
     main = name)
}
plotmodel(modelA_driv, 'modelA with drivers')
```
```{r}
plotmodel(modelA_driv2, 'modelA with drivers, but no propagules')
```

### 2) Marginal likelihood of the model under different scenarios (probability of generating a expected observations given the model, i.e. model evidence)

#### Scenario 1: High sediment supply
Expect seaward mangroves to increase

#### Scenario 2: High sediment supply and sea-level fall (e.g., during drought)
Expect Seaward mangroves to increase (progradation) and landward mangroves dieback

#### Scenario 3: High sediment supply and sea-level rise
Expect seaward mangroves to prograde, landward mangroves to migrate landward

#### Scenario 4: Sea-level rise
Expect landward mangroves to increase, seaward mangroves to decrease

```{r}

numsims <- 10000
likelihoods <- c()

simA_SLR <- system.sim_press_val(numsims, modelA_driv, perturb=c(SeaLevelRise=1), 
                                             monitor=c(SeawardMang=-1, LandwardMang=1))
marg_likelihood_SLR <- simA_SLR$accepted/simA_SLR$total
marg_likelihood_SLR
likelihoods[1] <- marg_likelihood_SLR

simA_SLRa <- system.sim_press_val(numsims, modelA_driv2, perturb=c(SeaLevelRise=1), 
                                             monitor=c(SeawardMang=-1, LandwardMang=1))
marg_likelihood_SLRa <- simA_SLRa$accepted/simA_SLRa$total
marg_likelihood_SLRa
likelihoods[5] <- marg_likelihood_SLRa

simA_SLR_SedSupp <- system.sim_press_val(numsims, modelA_driv,perturb=c(SeaLevelRise=1,
                                                                        SedSupply=1), 
                                             monitor=c(SeawardMang=1, LandwardMang=1))
marg_likelihood_SLR_SedSupp <- simA_SLR_SedSupp$accepted/simA_SLR_SedSupp$total
marg_likelihood_SLR_SedSupp
likelihoods[2] <- marg_likelihood_SLR_SedSupp

simA_SLR_SedSuppa <- system.sim_press_val(numsims, modelA_driv2,perturb=c(SeaLevelRise=1,
                                                                        SedSupply=1), 
                                             monitor=c(SeawardMang=1, LandwardMang=1))
marg_likelihood_SLR_SedSuppa <- simA_SLR_SedSuppa$accepted/simA_SLR_SedSuppa$total
marg_likelihood_SLR_SedSuppa
likelihoods[6] <- marg_likelihood_SLR_SedSuppa

simA_SLF_SedSupp <- system.sim_press_val(numsims, modelA_driv,perturb=c(SeaLevelFall=1,
                                                                        SedSupply=1), 
                                             monitor=c(SeawardMang=1, LandwardMang=-1))
marg_likelihood_SLF_SedSupp <- simA_SLF_SedSupp$accepted/simA_SLF_SedSupp$total
marg_likelihood_SLF_SedSupp
likelihoods[3] <- marg_likelihood_SLF_SedSupp

simA_SLF_SedSuppa <- system.sim_press_val(numsims, modelA_driv2,perturb=c(SeaLevelFall=1,
                                                                        SedSupply=1), 
                                             monitor=c(SeawardMang=1, LandwardMang=-1))
marg_likelihood_SLF_SedSuppa <- simA_SLF_SedSuppa$accepted/simA_SLF_SedSuppa$total
marg_likelihood_SLF_SedSuppa
likelihoods[7] <- marg_likelihood_SLF_SedSuppa

simA_SedSupp <- system.sim_press_val(numsims, modelA_driv,perturb=c(SedSupply=1), 
                                             monitor=c(SeawardMang=1))
marg_likelihood_SedSupp <- simA_SedSupp$accepted/simA_SedSupp$total
marg_likelihood_SedSupp
likelihoods[4] <- marg_likelihood_SedSupp

simA_SedSuppa <- system.sim_press_val(numsims, modelA_driv2,perturb=c(SedSupply=1), 
                                             monitor=c(SeawardMang=1))
marg_likelihood_SedSuppa <- simA_SedSuppa$accepted/simA_SedSuppa$total
marg_likelihood_SedSuppa
likelihoods[8] <- marg_likelihood_SedSuppa

datplot <- data.frame(scenario = rep(c('SLR', 'SLR & Sediment supply', 'SLF & Sediment supply', 'Sediment supply'),2),
                      model = c(rep('Driver model', 4), rep('Driver model simplified', 4)),
                      likelihood = likelihoods)

ggplot(datplot) +
  aes(x = scenario, y = likelihoods) +
  geom_bar(stat = 'identity') +
  facet_wrap(~model) +
  ylim(c(0,1)) +
  theme(axis.text.x = element_text(angle = 20,
                                   hjust = 1)) +
  ylab("Model evidence") +
  xlab("Scenario") +
  geom_hline(yintercept = 0.5)
```
Model evidence is generally low for both models. Note that simpler models without driver variables have been tested, and tend to have higher model evidence, but don't allow us to examine the responses of mangroves to multiple, interacting pressures, i.e., we cannot consider the effect of multiple drivers on a single variable of interest, i.e., mangrove increase/decrease. So we are using the more complex, driver based models.

We will use the simplified driver model for further investigation. Only if we want to examine tidal amplitude/tidal frequency scenarios should we use the more complex driver model with propagules.

### 3) Predicted response of mangroves under Kerrylee's 'validation' scenarios

Here we only consider models that are stable to estimate the probability of different mangrove responses (evaluation of model evidence above includes all models that are simulated (stable + unstable))

Blue is the probability of mangrove landward migration under each scenario.
Red is the probability of mangrove seaward migration (i.e., progradation) under each scenario.

```{r}

outcomes <- data.frame(rbind(simA_SLRa$allout, simA_SedSuppa$allout, simA_SLR_SedSuppa$allout, simA_SLF_SedSuppa$allout))

outcomes$scnr <- rep(c("SLR", "Sediment supply", "Sediment supply & SLR", "Sediment supply & SLF"), times = c(simA_SLRa$stable, simA_SedSuppa$stable, simA_SLR_SedSuppa$stable, simA_SLF_SedSuppa$stable))

outcomes <- outcomes %>% pivot_longer(LandwardLatAccom:SubVol, 
                         names_to = "variable", values_to = 'outcome')

# calculate proportion of stable models that have positive, negative, or neutral outcome in landward/seaward mangrove response

seaward <- outcomes %>% 
  filter(variable == 'SeawardMang') %>% 
  group_by(scnr) %>% 
  summarise(Increase = sum(outcome>0)/n(),
            Neutral = sum(outcome==0)/n(),
            Decrease = sum(outcome<0)/n()) %>% 
  pivot_longer(Increase:Decrease ,names_to = 'outcome', values_to = 'prop') %>% 
  mutate(outcome = factor(outcome, levels = c('Increase', 'Neutral', 'Decrease')))

landward <- outcomes %>% 
  filter(variable == 'LandwardMang') %>% 
  group_by(scnr) %>% 
    summarise(Increase = sum(outcome>0)/n(),
            Neutral = sum(outcome==0)/n(),
            Decrease = sum(outcome<0)/n()) %>% 
  pivot_longer(Increase:Decrease ,names_to = 'outcome', values_to = 'prop') %>% 
  mutate(outcome = factor(outcome, levels = c('Increase', 'Neutral', 'Decrease')))

# Note, as potential TODO could boostrap resample here to get estimate of uncertainty around that probability

a <- ggplot(seaward) +
  geom_bar(aes(y = scnr, x = prop, fill = outcome),
           position = 'stack', stat = 'identity') +
  xlab('Proportion of outcomes') +
  ylab('') +
  theme(legend.position = 'none') +
  ggtitle('Seaward mangroves')

b <- ggplot(landward) +
  geom_bar(aes(y = scnr, x = prop, fill = outcome),
           position = 'stack', stat = 'identity') +
  xlab('Proportion of outcomes') +
  ylab('') +
  ggtitle('Landward mangroves') +
  theme(legend.title = element_blank(),
        axis.text.y =  element_blank())

a+b
```
Landward mangroves always do what we expect under each scenario, except for scenario of sediment supply - under which we have no expectation of how landward mangroves will respond (we only expect progradation of mangroves). 

The response of seaward mangroves is more uncertain, except for scenario of sea-level rise where we see progradation as expected. For the others, seaward mangroves do what we expect most of the time.

Eventually we can map these as qualitative predictions about the likely response of mangroves given different drivers.

### 4) For each scenario, get models that are valid, construct probability distribiution of likely parameter values

Quantify and compare uncertainty in parameters. Where are valid parameters most uncertain? This suggest where we should focus our attention to resolve uncertainty in parameters (e.g., collect more/better data) to make better predictions.

Parameters are taken from simulated models that meet the following criteria: a) stable, and b) are valid under one or all scenarios.

```{r}

params <- data.frame(rbind(simA_SLRa$valweights, simA_SedSuppa$valweights, simA_SLR_SedSuppa$valweights, simA_SLF_SedSuppa$valweights))

params$scnr <- rep(c("SLR", "Sediment supply", "Sediment supply & SLR", "Sediment supply & SLF"), each = numsims)

params <- params %>% pivot_longer(LandwardLatAccom....LandwardMang:TidalFreq....TidalFreq, 
                         names_to = "link", values_to = 'weight')

params$from_var <- sapply(strsplit(as.character(params$link), "\\."), `[`, 1)
params$to_var <- sapply(strsplit(as.character(params$link), "\\."), `[`, 5)

paramsvis <- subset(params, from_var != to_var)

ggplot() +
  geom_density(data = filter(paramsvis, scnr == 'SLR'), 
               aes(x = weight), colour = 'blue', fill = 'blue', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == 'Sediment supply'), 
               aes(x = weight), colour = 'yellow', fill = 'yellow', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == 'Sediment supply & SLR'), 
               aes(x = weight), colour = 'green', fill = 'green', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == 'Sediment supply & SLF'), 
               aes(x = weight), colour = 'red', fill = 'red', alpha = 0.2) +
  # geom_density(data = filter(params, scnr == 'All'), 
   #            aes(x = weight), colour = 'purple', fill = 'purple', alpha = 0.2) +
  facet_wrap(~link, scales = 'free') +
  theme(strip.text.x = element_text(size = 5))

```
The scenarios are valid under a generally very ambiguous set of parameter weights. 
That probably means our predictions are very uncertain too.

TODO: Try to put tighter constraints on parameter values with more scenario validation
Make this into a function and put in an app, so that you can easily select for very complicated press perturbation scenarios for all drivers, and see outcome.

1) Do more interesting scenarios with lots of drivers
2) Determine the relative importance of those drivers to the outcome (+ve, -ve, neutral). How to do this when drivers aren't in the model, i.e., can't just see how parameter weights relate to the predicted outcome. 
- perhaps frame as the relative importance of drivers to the uncertainty in linkage weights, and therefore model predictions? If do this way, can evaluate by running a model with categorical predi

### 5) SLR scenarios with other climate and anthropogenic impacts, and conservation/restoration actions

#### Scenario 1: SLR and cyclones
Expect mangroves to migrate landwards and seaward mangroves to decrease

#### Scenario 2: SLR and cyclones and dams

#### Scenario 3: SLR and cyclones and Dams and coastal development

#### Scenario 4: SLR and cyclones and Dams and coastal development and erosion

#### Scenario 5: SLR and cyclones and Dams and coastal development and active transplantation

#### Scenario 6: SLR and cyclones and Dams and coastal development and active transplantation and permeable dams

```{r}

simA_SLR_cyc <- system.sim_press_val(numsims, modelA_driv2, perturb=c(SeaLevelRise=1,
                                                               Cyclones=1), 
                                             monitor=c(SeawardMang=-1))

simA_SLR_cyc_dams <- system.sim_press_val(numsims, modelA_driv2, perturb=c(SeaLevelRise=1,
                                                               Cyclones=1,
                                                               Dams=1), 
                                             monitor=c(SeawardMang=-1))

simA_SLR_cyc_dams_cd <- system.sim_press_val(numsims, modelA_driv2, perturb=c(SeaLevelRise=1,
                                                               Cyclones=1,
                                                               Dams=1,
                                                               CoastalDev=1), 
                                             monitor=c(SeawardMang=-1))

simA_SLR_cyc_dams_cd_rest <- system.sim_press_val(numsims, modelA_driv2, perturb=c(SeaLevelRise=1,
                                                               Cyclones=1,
                                                               Dams=1,
                                                               CoastalDev=1, 
                                                               Restoration=1), 
                                             monitor=c(SeawardMang=-1))

simA_all <- system.sim_press_val(numsims, modelA_driv2, perturb=c(SeaLevelRise=1,
                                                               Cyclones=1,
                                                               Dams=1,
                                                               CoastalDev=1, 
                                                               LandwardMang=1,
                                                               SeawardMang=1), 
                                             monitor=c(SeawardMang=-1))


outcomes <- data.frame(rbind(simA_SLR_cyc$allout, simA_SLR_cyc_dams$allout, simA_SLR_cyc_dams_cd$allout, simA_SLR_cyc_dams_cd_rest$allout))

outcomes$scnr <- rep(c("SLR & cyclones", "SLR & cyclones & dams", "SLR & cyclones & dams & coastal dev", "SLR & cyclones & dams & coastal dev & restoration"), times = c(simA_SLR_cyc$stable-1, simA_SLR_cyc_dams$stable-1, simA_SLR_cyc_dams_cd$stable-1, simA_SLR_cyc_dams_cd_rest$stable-1))

outcomes <- outcomes %>% pivot_longer(colnames(outcomes[1]):colnames(outcomes[ncol(outcomes)-1]), 
                         names_to = "variable", values_to = 'outcome')

# calculate proportion of stable models that have positive, negative, or neutral outcome in landward/seaward mangrove response

seaward <- outcomes %>% 
  filter(variable == 'SeawardMang') %>% 
  group_by(scnr) %>% 
  summarise(Increase = sum(outcome>0)/n(),
            Neutral = sum(outcome==0)/n(),
            Decrease = sum(outcome<0)/n()) %>% 
  pivot_longer(Increase:Decrease ,names_to = 'outcome', values_to = 'prop') %>% 
  mutate(outcome = factor(outcome, levels = c('Increase', 'Neutral', 'Decrease')))

landward <- outcomes %>% 
  filter(variable == 'LandwardMang') %>% 
  group_by(scnr) %>% 
    summarise(Increase = sum(outcome>0)/n(),
            Neutral = sum(outcome==0)/n(),
            Decrease = sum(outcome<0)/n()) %>% 
  pivot_longer(Increase:Decrease ,names_to = 'outcome', values_to = 'prop') %>% 
  mutate(outcome = factor(outcome, levels = c('Increase', 'Neutral', 'Decrease')))

# Note, as potential TODO could boostrap resample here to get estimate of uncertainty around that probability

a <- ggplot(seaward) +
  geom_bar(aes(y = scnr, x = prop, fill = outcome),
           position = 'stack', stat = 'identity') +
  xlab('Proportion of outcomes') +
  ylab('') +
  theme(legend.position = 'none') +
  ggtitle('Seaward mangroves')

b <- ggplot(landward) +
  geom_bar(aes(y = scnr, x = prop, fill = outcome),
           position = 'stack', stat = 'identity') +
  xlab('Proportion of outcomes') +
  ylab('') +
  ggtitle('Landward mangroves') +
  theme(legend.title = element_blank(),
        axis.text.y =  element_blank())

a+b
```

```{r}

params <- data.frame(rbind(simA_SLR_cyc$valweights, simA_SLR_cyc_dams$valweights, simA_SLR_cyc_dams_cd$valweights, simA_SLR_cyc_dams_cd_rest$valweights))

params$scnr <- rep(c("SLR & cyclones", "SLR & cyclones & dams", "SLR & cyclones & dams & coastal dev", "SLR & cyclones & dams & coastal dev & restoration"), each = numsims)

params <- params %>% pivot_longer(colnames(params[1]):colnames(params[ncol(params)-1]),
                         names_to = "link", values_to = 'weight')

params$from_var <- sapply(strsplit(as.character(params$link), "\\."), `[`, 1)
params$to_var <- sapply(strsplit(as.character(params$link), "\\."), `[`, 5)

paramsvis <- subset(params, from_var != to_var)

ggplot() +
  geom_density(data = filter(paramsvis, scnr == 'SLR'), 
               aes(x = weight), colour = 'blue', fill = 'blue', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == 'SLR & cyclones'), 
               aes(x = weight), colour = 'yellow', fill = 'yellow', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == 'SLR & cyclones & dams'), 
               aes(x = weight), colour = 'green', fill = 'green', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == 'SLR & cyclones & dams & coastal dev'), 
               aes(x = weight), colour = 'red', fill = 'red', alpha = 0.2) +
   geom_density(data = filter(paramsvis, scnr == "SLR & cyclones & dams & coastal dev & restoration"), 
               aes(x = weight), colour = 'purple', fill = 'purple', alpha = 0.2) +
  facet_wrap(~link, scales = 'free') +
  theme(strip.text.x = element_text(size = 5))

```





