---
output: html_document
---

TODO: Create many scenarios to try and test. Figure out whats next. Any uncertain linkages?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(QPress)
library(dplyr)
library(ggplot2)
theme_set(theme_classic())
```

# How to build a model 
### Mangrove response to climate change and other anthropogenic threats

Taking a 'one model to rule them all approach' vs. multiple, individual models for different scenarios. We agree there should be only one model structure. If we think the model structure isn't right though, we can test and compare other variations.

Aim is to build one model that performs as expected under different scenarios.

Then we can map likelihood of increase or decrease in mangrove extent, given different drivers

Steps are to:

1) Build and visualise the conceptual model
2) Perturb drivers and simulate to examine expected responses of mangroves under different scenarios
3) If responses are not expected, re-work the conceptual model and do step 2
4) Map likelihood of mangroves increasing or decreasing given different geomorphologies and drivers
5) Validate with GMW data (if possible)

### 1. Build and visualise conceptual model

Variables are in white and drivers are in grey. Red arrows indicate negative effect, black arrows indicate positive effect.

***NOTE currently running model assuming back-front are stable. Should try alternate model structure and compare.

```{r}
base_model <- c('LandwardLatAccom->LandwardPropag',
                'LandwardPropag->LandwardMang',
                'LandwardMang-->OM',
                #'LandwardMang->OM',
                #'LandwardMang-*LandwardMang',
                #'LandwardPropag-*LandwardPropag',
                'SeawardLatAccom->SeawardPropag',
                'SeawardPropag->SeawardMang',
                'SeawardMang->OM',
                #'SeawardMang-*SeawardMang',
                #'SeawardPropag-*SeawardPropag',
                #'LandwardLatAccom-*SeawardLatAccom',
                #'SeawardLatAccom-*LandwardLatAccom',
                #'LandwardPropag-*SeawardPropag',
                #'SeawardPropag-*LandwardPropag',
                #'LandwardMang-*SeawardMang',
                #'SeawardMang-*LandwardMang',
                'SeaLevelFall-*LandwardLatAccom',
                'SeaLevelFall--*LandwardMang',
                'SeaLevelFall->SeawardLatAccom',
                'SeaLevelRise->LandwardLatAccom',
                'SeaLevelRise--*SeawardMang',
                'SeaLevelRise->SeawardLatAccom',
                'Cyclones--*LandwardMang',
                'Cyclones->SeawardMang',
                'Restoration->LandwardPropag',
                'Restoration->SeawardPropag',
                'TidalAmp->LandwardLatAccom',
                'TidalAmp->SeawardLatAccom',
                'BasementUp-*LandwardLatAccom',
                'BasementUp--*LandwardMang',
                'BasementUp->SeawardLatAccom',
                'Subsidence->LandwardLatAccom',
                'Subsidence--*SeawardMang',
                'Subsidence-*SeawardLatAccom',
                'CoastalDev-*LandwardLatAccom',
                'CoastalDev--*LandwardMang',
                'CoastalDev--*SeawardMang',
                'Protection-*CoastalDev',
                'Erosion-*SeawardMang',
                'PermeableDam-*Erosion',
                'HydroEnergy->Erosion',
                'OM->SubVol',
                'TidalFreq-*SeawardPropag',
                'TidalFreq-*SubVol',
                'SubVol->SeawardLatAccom',
                #'SubVol-*SubVol'
                'Autocompaction-*SubVol',
                'SedSupply-*SubVol',
                'Dams-*SedSupply',
                'HydroEnergy-*SeawardLatAccom',
                'HydroEnergy-*SeawardPropag'
                )
                
drivers <- c("Cyclones","SeaLevelFall","Restoration","CoastalDev",
             "SeaLevelRise","TidalAmp", "TidalFreq", "Autocompaction", "SedSupply", 
             "Dams", "Protection", "PermeableDam", "Erosion", "HydroEnergy",
             "Subsidence", "BasementUp")
modelA <- parse.digraph(base_model)
modelA <- enforce.limitation(modelA)
```


```{r dpi=300, fig.width=7, fig.height=7}
plotmodel <- function(modelA){
w <- adjacency.matrix(modelA, labels=TRUE)
gdat <- graph_from_adjacency_matrix(t(w),
                                    mode = "directed",
                                    weighted = "b")
E(gdat)$color <- ifelse(E(gdat)$b <0, "red", "black")
V(gdat)$color <- ifelse(names(V(gdat)) %in% drivers, "grey", "white")
plot(gdat, 
     vertex.size = 10,
     vertex.label.cex = 0.5,
     vertex.label.color = "black",
     edge.arrow.size = 0.3,
     #edge.label = letters[1:length(E(gdat))],
     edge.label.cex = 0.6,
     edge.label.color = "black",
     edge.label.pos = 1)
}
plotmodel(modelA)
```

### 2) Set-up scenarios

#### Scenario 1: High sediment supply

Expect seaward mangroves to increase

```{r}
#sample community
s <- community.sampler(modelA) # constructs functions to generate random community matrices, provides a list of functions
s$select(0.5) # randomly select which uncertain pairs will be retained with a probability of 0.5
W <- s$community() # generate a random community matrix

# this is monitoring just a single community matrix
# perturb sediment supply, monitor all
impactSedSupply <- press.impact(modelA, perturb = c(SedSupply=1))
results <- round(impactSedSupply(W),2)
names(results) <- colnames(adjacency.matrix(modelA, labels = T))
results

# perturb sediment supply, and see if seaward mangroves increase
valSedSupply <- press.validate(modelA, perturb=c(SedSupply=1), monitor=c(SeawardMang=1))
valSedSupply(W)
```

#### Scenario 2: High sediment supply and sea-level fall (e.g., during drought)

Expect Seaward mangroves to increase (progradation) and landward mangroves dieback

```{r}
impactSedSupply_SLF <- press.impact(modelA, perturb = c(SedSupply=1, 
                                                        SeaLevelFall=1))
results <- round(impactSedSupply_SLF(W),2)
names(results) <- colnames(adjacency.matrix(modelA, labels = T))
results

# perturb sediment supply, and see if seaward mangroves increase
valSedSupply_SLF <- press.validate(modelA, perturb=c(SedSupply=1,
                                                     SeaLevelFall=1), 
                                   monitor=c(SeawardMang=1, LandwardMang=-1))
valSedSupply_SLF(W)
```

#### Scenario 3: High sediment supply and sea-level rise

Expect seaward mangroves to prograde, landward mangroves to migrate landward

```{r}
impactSedSupply_SLR <- press.impact(modelA, perturb = c(SedSupply=1, 
                                                        SeaLevelRise=1))
results <- round(impactSedSupply_SLR(W),2)
names(results) <- colnames(adjacency.matrix(modelA, labels = T))
results

# perturb sediment supply, and see if seaward mangroves increase
valSedSupply_SLR <- press.validate(modelA, perturb=c(SedSupply=1,
                                                     SeaLevelRise=1), 
                                   monitor=c(SeawardMang=1, LandwardMang=1))
valSedSupply_SLR(W)
```

#### Scenario 4: Sea-level rise

Expect landward mangroves to increase, seaward mangroves to decrease

```{r}
impact_SLR <- press.impact(modelA, perturb = c(SeaLevelRise=1))
results <- round(impactSedSupply_SLR(W),2)
names(results) <- colnames(adjacency.matrix(modelA, labels = T))
results

# perturb sediment supply, and see if seaward mangroves increase
val_SLR <- press.validate(modelA, perturb=c(SeaLevelRise=1), 
                                   monitor=c(SeawardMang=-1, LandwardMang=1))
val_SLR(W)
```

### 3) Simulate scenarios and compare outcomes

```{r}
simulate_scenarios <- function(i){
  
  simB <- community.sampler(modelA)
  simB$select(0.5) # randomly select which uncertain pairs will be retained with a probability of 0.5
  varnames <- colnames(adjacency.matrix(modelA, labels = T))
  
  stable_comm <- FALSE
  
  while(!stable_comm){
    w <- simB$community()
    stable_comm <- stable.community(w)
  }
  
  modout <- data.frame(isim = i,
                       scnr = c("Sediment supply", "Sea-level rise", 
                                    "Sediment supply & Sea-level fall",
                                    "Sediment supply & Sea-level rise"),
                       outcomes = c(valSedSupply(w),
                                    val_SLR(w),
                                    valSedSupply_SLF(w),
                                    valSedSupply_SLR(w))
  )
  return(modout)
}
modout <- lapply(1:1000, simulate_scenarios)
modout <- do.call("rbind", modout)
```

Below graph shows proportion of simulations where model is valid across 1000 simulations with randomised weights, with or without uncertain linkages (latter is TODO)

```{r}
#summarise results
modsum <- modout %>% group_by(scnr) %>%
  summarize(prop_valid = sum(outcomes==TRUE)/n())
ggplot(modsum) +
  aes(x = scnr, y = prop_valid) + 
  geom_bar(stat = 'identity') +
  ylim(c(0,1)) +
  theme(axis.text.x = element_text(angle = 20,
                                   hjust = 1)) +
  ylab("Proportion valid") +
  xlab("Scenario") +
  geom_hline(yintercept = 0.5)
```


```{r}
mang_dens <- modout %>% filter(vars == "LandwardMang")
ggplot(mang_dens) +
   aes(x = outcomes) + 
  geom_density(color = "red", fill = "pink") +
  facet_wrap(~scnr, scales= "free") + 
  xlim(-10, 10) + 
  xlab("Net outcome") +
  geom_vline(xintercept = 0)
```


```{r}
mang_dens <- modout %>% filter(vars == "SeawardMang")
ggplot(mang_dens) +
   aes(x = outcomes) + 
  geom_density(color = "red", fill = "pink") +
  facet_wrap(~scnr, scales= "free") + 
  xlim(-10, 10) + 
  xlab("Net outcome") +
  geom_vline(xintercept = 0)
```


